Drivers of sex differences in consumption (NAMCS)

### 2017-07-05 sex-difference/sex-medicare-display.Rmd
initial

```{r load_data, cache=FALSE}
sum_over = function(target_columns) {
  dummy_vars = sapply(1:length(target_columns), function(i) sprintf('x%i', i))
  dummy_vars_string = paste0(dummy_vars, collapse=',')
  dummy_vars_list = lapply(target_columns, as.name) %>% setNames(dummy_vars)
  f = as.formula(paste0(c('~sum(', dummy_vars_string, ')'), collapse=''))
  lazyeval::interp(f, .values=dummy_vars_list)
}

regions = read_tsv('../../db/census-regions/census-regions.tsv') %>%
  select(state, region)

condition_names = read_tsv('../chronic-conditions/names.tsv') %>% filter(condition_code != 'ALZH')
names_1yr_ccs = condition_names %>% filter(condition_ref_years==1) %$% condition_code

dx_codes = read_tsv('../sex-difference/sex_codes.tsv')

load_data = function(year) {
  # pde data
  pde = read_tsv(sprintf('../pde_%i.tsv', year)) %>%
    select(bene_id, antibiotic, days_supply) %>%
    group_by(bene_id, antibiotic) %>%
    summarize(n_claims=n(), days_supply=sum(days_supply)) %>%
    ungroup() %>%
    mutate(year=year)

  # diagnosis claims
  dx_car_fn = sprintf('../../data/dx_car_claims_%i.tsv', year)
  dx_op_fn = sprintf('../../data/dx_op_claims_%i.tsv', year)

  dx = bind_rows(read_tsv(dx_car_fn),
                 read_tsv(dx_op_fn)) %>%
    mutate(date=dmy(from_date)) %>%
    rename(bene_id=BENE_ID, code=diagnosis) %>%
    left_join(dx_codes, by='code')

  dx_presence = dx %>%
    group_by(bene_id) %>%
    summarize(rc='acute_rc' %in% diagnosis_type,
              uti='uti' %in% diagnosis_type)

  count_car_fn = sprintf('../../data/count_car_claims_%i.tsv', year)
  count_op_fn = sprintf('../../data/count_op_claims_%i.tsv', year)
  n_op_days = bind_rows(read_tsv(count_car_fn), read_tsv(count_op_fn)) %>%
    rename(bene_id=BENE_ID) %>%
    group_by(bene_id) %>%
    summarize(n_op_days=sum(n_from_date))

  # total numbers of PDEs
  n_claims = pde %>% group_by(bene_id) %>% summarize(n_claims=sum(n_claims))

  # count chronic conditions. 'n_cc' means one-year cc's
  cc = read_feather(sprintf('../cc_%i.feather', year)) %>%
    rowwise() %>%
    mutate_(n_cc=sum_over(names_1yr_ccs)) %>%
    ungroup()

  # join the cc, summary PDE, and dx data into bene
  bene = read_tsv(sprintf('../bene_%i.tsv', year)) %>%
    mutate(is_female=sex=='female', is_white=race=='white', is_dual=buyin_months>0) %>%
    left_join(regions, by='state') %>%
    mutate(year=year) %>%
    left_join(cc, by='bene_id') %>%
    left_join(n_op_days, by='bene_id') %>% replace_na(list(n_op_days=0)) %>%
    left_join(n_claims, by='bene_id') %>% replace_na(list(n_claims=0)) %>%
    left_join(dx_presence, by='bene_id') %>% replace_na(list(rc=FALSE, uti=FALSE))

  list(bene=bene, pde=pde, dx=dx)
}

years = 2011:2014
n_years = length(years)
dat = lapply(years, load_data)

bene = lapply(dat, function(df) df$bene) %>% bind_rows %>%
  group_by(bene_id) %>%
  mutate(n_years=n(),
         start_age=min(age),
         in_cohort=n_years==4) %>%
  ungroup()

pde = lapply(dat, function(df) df$pde) %>% bind_rows %>%
  left_join(distinct(select(bene, bene_id, in_cohort)), by='bene_id')

dx = lapply(dat, function(df) df$dx) %>% bind_rows() %>%
  left_join(distinct(select(bene, bene_id, in_cohort)), by='bene_id')
```

@@

```{r dx_counts}
dx %>%
  filter(in_cohort) %>%
  count(code) %>%
  left_join(dx_codes, by='code') %>%
  arrange(diagnosis_type, desc(n)) %>%
  mutate(n=format(n, big.mark=',')) %>%
  kable()
```

@@

```{r top_drugs}
top_abx = pde %>%
  count(antibiotic) %>%
  arrange(desc(n)) %>%
  head(10) %T>%
  kable %$%
  antibiotic
```

@@

```{r drugs_by_sex}
usage_by_sex = function(drug) {
  bene %>%
    filter(in_cohort) %>%
    left_join(pde %>% filter(antibiotic==drug) %>% count(bene_id), by='bene_id') %>%
    replace_na(list(n=0)) %>%
    group_by(sex) %>%
    summarize(n_bene=n(),
              n_using=sum(n>0),
              n_claims=sum(n),
              mean_n_claims=mean(n),
              sd_n_claims=sd(n)) %>%
    mutate(se_n_claims=sd_n_claims/sqrt(n_bene),
           drug=drug)
}

ubs = lapply(top_abx, usage_by_sex) %>% bind_rows
```

```{r healthcare_usage_by_sex}
hc_by_sex = dx %>%
  count(bene_id, diagnosis_type) %>%
  ungroup() %>%
  right_join(bene, by='bene_id') %>%
  replace_na(list(n=0)) %>%
  group_by(sex) %>%
  summarize(total_n_days=sum(n),
            mean_n_days=mean(n),
            n_using=sum(n>0),
            n_bene=n())

hc_use_ratio = hc_by_sex %>% mutate(ratio=mean_n_days/min(mean_n_days)) %$% max(ratio)
```

@@

```{r cipro}
cipro_dat = pde %>%
  filter(in_cohort, antibiotic=='ciprofloxacin') %>%
  select(bene_id, year, cipro=n_claims) %>%
  right_join(filter(bene, in_cohort), by=c('bene_id', 'year')) %>%
  replace_na(list(cipro=0)) %>%
  group_by(bene_id) %>%
  summarize(is_female=any(is_female),
            uti=any(uti),
            cipro=sum(cipro))

model_f = function(df, f) glm(f, data=df, family='poisson') %>% tidy

cipro_results = bind_rows(model_f(cipro_dat, cipro ~ is_female) %>% mutate(model='base'),
                          model_f(cipro_dat, cipro ~ is_female + uti) %>% mutate(model='expl'))
```

```{r cipro_results}
# fraction of risk ratio explained
frre = function(res, tm='is_femaleTRUE') {
  rr0 = res %>% filter(model=='base', term==tm) %$% exp(estimate)
  rr1 = res %>% filter(model=='expl', term==tm) %$% exp(estimate)
  (rr0 - rr1) / (rr0 - 1)
}

@@

```{r bactrim}
bactrim_dat = pde %>%
  filter(in_cohort, antibiotic=='trimethoprim/sulfamethoxazole') %>%
  select(bene_id, year, bactrim=n_claims) %>%
  right_join(filter(bene, in_cohort), by=c('bene_id', 'year')) %>%
  replace_na(list(bactrim=0)) %>%
  group_by(bene_id) %>%
  summarize(is_female=any(is_female),
            uti=any(uti),
            bactrim=sum(bactrim))

bactrim_results = bind_rows(model_f(bactrim_dat, bactrim ~ is_female) %>% mutate(model='base'),
                            model_f(bactrim_dat, bactrim ~ is_female + uti) %>% mutate(model='expl'))

bactrim_results %>% kable
```

@@

```{r nitro}
nitro_dat = pde %>%
  filter(in_cohort, antibiotic=='nitrofurantoin') %>%
  select(bene_id, year, n=n_claims) %>%
  right_join(filter(bene, in_cohort), by=c('bene_id', 'year')) %>%
  replace_na(list(n=0)) %>%
  group_by(bene_id) %>%
  summarize(is_female=any(is_female),
            uti=any(uti),
            n=sum(n))

nitro_results = bind_rows(model_f(nitro_dat, n ~ is_female) %>% mutate(model='base'),
                          model_f(nitro_dat, n ~ is_female + uti) %>% mutate(model='expl'))

nitro_results %>% kable
```

@@

```{r women_azithro_vs_uai, eval=FALSE}
azithro_dat = pde %>%
  filter(in_cohort, antibiotic=='azithromycin') %>%
  select(bene_id, year, n=n_claims) %>%
  right_join(filter(bene, in_cohort), by=c('bene_id', 'year')) %>%
  replace_na(list(n=0)) %>%
  group_by(bene_id) %>%
  summarize(is_female=any(is_female),
            rc=any(rc),
            copd=any(COPD),
            n=sum(n))

azithro_results = bind_rows(model_f(azithro_dat, n ~ is_female) %>% mutate(model='base'),
                            model_f(azithro_dat, n ~ is_female + rc + copd) %>% mutate(model='expl'))

azithro_results %>% kable
```

@@

```{r consumption_by_age_and_sex}
bene %>%
  filter(in_cohort) %>%
  group_by(bene_id) %>%
  summarize(is_female=any(is_female),
            n=sum(n_claims),
            age=min(age)) %>%
  ggplot(aes(x=age, y=n, group=is_female)) +
  geom_boxplot()
```

@@

```{r sex_by_age}
bene %>% filter(in_cohort) %>%
  ggplot(aes(x=age, fill=sex)) +
  geom_bar(position='dodge')
```

@@

```{r azithro_even}
pde %>%
  filter(antibiotic=='azithromycin') %>%
  group_by(bene_id) %>% summarize(azi=sum(n_claims)) %>%
  right_join(bene %>% filter(in_cohort) %>% select(bene_id, sex) %>% distinct,
    by='bene_id') %>% replace_na(list(azi=0)) %>%
  count(sex, azi) %>%
  group_by(sex) %>%
  mutate(f_ppl=n/sum(n), cum_f_ppl=cumsum(f_ppl),
         cum_cons=cumsum(azi*n), cum_f_cons=cum_cons/max(cum_cons)) %>%
  ggplot(aes(x=cum_f_ppl, y=cum_f_cons, color=sex)) +
  geom_point() +
  geom_line() +
  xlab('Fraction of consumers') + ylab("Fraction of that sex's azithromycin consumed")
```

>>>>

```{r load_data}
years = 2011:2014
n_years = length(years)

dx_codes = read_tsv('sex_codes.tsv')
```

@@

```{r dx_counts}
read_tsv('data/tbl_dx_counts.tsv') %>%
  left_join(dx_codes, by='code') %>%
  arrange(diagnosis_type, desc(n)) %>%
  mutate(n=format(n, big.mark=',')) %>%
  kable()
```

@@

```{r top_drugs}
top_abx = read_tsv('data/tbl_top_abx.tsv') %>%
  arrange(desc(n)) %>%
  head(10) %T>%
  kable %$%
  antibiotic
```

@@

```{r usage_by_sex}
# drugs by sex
ubs = read_tsv('data/tbl_usage_by_sex.tsv')

# healthcare usage by sex
hc_by_sex = read_tsv('tbl_hc_by_sex.tsv')
hc_use_ratio = hc_by_sex %>% mutate(ratio=mean_n_days/min(mean_n_days)) %$% max(ratio)
```

@@

```{r cipro}
cipro_results = read_tsv('data/tbl_model_cipro.tsv')

# fraction of risk ratio explained
frre = function(res, tm='is_femaleTRUE') {
  rr0 = res %>% filter(model=='base', term==tm) %$% exp(estimate)
  rr1 = res %>% filter(model=='expl', term==tm) %$% exp(estimate)
  (rr0 - rr1) / (rr0 - 1)
}

@@

```{r bactrim}
bactrim_results = read_tsv('data/tbl_model_tmpsmx.tsv')
bactrim_results %>% kable
```

@@

```{r nitro}
nitro_results = read_tsv('data/tbl_model_nitro.tsv')
nitro_results %>% kable
```

@@

```{r women_azithro_vs_uai}
azithro_results = read_tsv('data/tbl_model_azithro.tsv')
azithro_results %>% kable
```

@@

```{r consumption_by_age_and_sex}
read_tsv('data/tbl_consumption_age_sex.tsv') %>%
  ggplot(aes(x=age, y=n, group=is_female)) +
  geom_boxplot()
```

@@

```{r sex_by_age}
read_tsv('data/tbl_sex_by_age.tsv') %>%
  ggplot(aes(x=age, y=n, fill=sex)) +
  geom_bar(stat='identity', position='dodge')
```

@@

```{r azithro_even}
read_tsv('data/tbl_azithro_evenness.tsv') %>%
  ggplot(aes(x=cum_f_ppl, y=cum_f_cons, color=sex)) +
  geom_point() +
  geom_line() +
  xlab('Fraction of consumers') + ylab("Fraction of that sex's azithromycin consumed")
```

### 2017-07-05 sex-difference/sex-namcs.Rmd
initial

dx_visit = dx %>%
  group_by(visit_id) %>%
  summarize(uti='uti' %in% diagnosis_type,
            acne='acne' %in% diagnosis_type)

@@

visits = read_tsv_years('../../db/namcs/data/namcs_visits_%i.tsv') %>%
  mutate(age_group=case_when(.$age <= 19 ~ 'child',
                             .$age <= 64 ~ 'adult',
                             .$age <= 92 ~ 'elderly')) %>%
  left_join(dx_visit, by='visit_id') %>%
  replace_na(list(uti=FALSE, acne=FALSE)) %>%
  left_join(n_abx, by='visit_id') %>% replace_na(list(n_abx=0)) %>%
  left_join(abx_visit, by='visit_id') %>%
  mutate_at(top_abx$abx, function(x) if_else(is.na(x), 0L, x))

@@

I kept the drugs from azithromycin through clarithromycin, accounting for 80% of all prescriptions.

@@

Overall, more of the visits were made by women. Women make `r round(female_visit_odds[2], 3)*100`% more visits than men. I'll therefore evaluate consumption by two standards: first, do women consume more antibiotics overall; second, do women consume more antibiotics *per visit*.

@@

@@

## Clindamycin

```{r clinda}
m1 = svyglm(clinda~sex, design=adult_design, family='poisson')
m2 = svyglm(clinda~sex+acne, design=adult_design, family='poisson')
clinda_f_expl = 1 - exp(-m2$coefficients['sexmale']+m1$coefficients['sexmale'])
```

The only clindamycin-associated diagnosis that meets (maybe) the data validity criterion is acne (`7061-`). However, acne explains only a small fraction (`r 100*round(clinda_f_expl, digits=3)`%) of the difference between the sexes.

>>>>

dx_visit = dx %>%
  group_by(visit_id) %>%
  summarize(uti='uti' %in% diagnosis_type,
            acne='acne' %in% diagnosis_type,
            arc='acute_rc' %in% diagnosis_type)

@@

visits = read_tsv_years('../../db/namcs/data/namcs_visits_%i.tsv') %>%
  mutate(age_group=case_when(.$age <= 19 ~ 'child',
                             .$age <= 64 ~ 'adult',
                             .$age <= 92 ~ 'elderly')) %>%
  left_join(dx_visit, by='visit_id') %>%
  replace_na(list(uti=FALSE, acne=FALSE, arc=FALSE)) %>%
  left_join(n_abx, by='visit_id') %>% replace_na(list(n_abx=0)) %>%
  left_join(abx_visit, by='visit_id') %>%
  mutate_at(top_abx$abx, function(x) if_else(is.na(x), 0L, x))

@@

I kept the drugs from azithromycin through cefdinir, but excluding clindamycin, accounting for 80% of all prescriptions. I exclude clindamycin because NAMCS doesn't have any information about route of administration. In some cases, the old IDs can be used to distinguish different form, but other times there is no hope. For example, in the 2011 data, there are 118 visits that include clindamycin. In 84 cases, it is just "clindamycin"; in 18, it is "Cleocin" (which can be oral, injected, or vaginal); in 9 cases, it is "clindamycin phosphate", which is normally topical, etc.

@@

Overall, more of the visits were made by women. Women make `r round(female_visit_odds[2]-1, 3)*100`% more visits than men.

What fraction of these visits are due to UTIs? We can see what fraction of visits are made for UTIs:

```{r}
svytable(~sex+uti, adult_design) %>%
  as_data_frame() %>%
  mutate(n=mpy(n)) %>%
  kable
```

So of the ~150 million extra visits per year made by women, a difference of ~5 million visits (or 3%) can be attributed to UTIs.

Of the top 3 diagnoses (hypertension, gynecological exam, and observation of pregnancy), two are woman-specific. Those account for 25 million visits per year, or about one-sixth of the total. So the difference is diffuse across diagnoses. Very few antibiotics are prescribed during those visits, less than 1 million per year.

Acute respiration complaints explain another ~5 million:

```{r}
svytable(~sex+arc, adult_design) %>%
  as_data_frame() %>%
  mutate(n=mpy(n)) %>%
  kable
```

@@

I'll therefore evaluate consumption by two standards: first, do women consume more antibiotics overall; second, do women consume more antibiotics *per visit*.

@@
